Modelling single-cell dynamics\\with trajectories and gene regulatory networks

### 2019-11-23 rebuttal/rebuttal.tex

\exam[\tagimp]{\imp{Important matter} $\rightarrow$}

\exam[\tagtime]{Matter which will take some time to reply to / fix $\rightarrow$}

\answ{Response or statement by candidate.}

\section{Dries De Maeyer}

\exam{The scope of the complete thesis is to define state transitions in cell types and the
	definition of cell type/state transition is not clear cut. I would frame the definition of a cell
	type more based on our current understanding and technology of biology. I.e. where do you
	draw the line of a new cell type?}

\exam{Could you make the main outline of the thesis could even be more clear by
	providing a graphical overview of the different topics addressed in this thesis?}

\subsection{Chapter 2}
\exam{Would you be able to extend this method to include Citeseq data? How would
	you see this?}

\exam{Can you use the generated models for validation with existing cellular
	heterogeneity from scRNA-seq data?}

\exam{Since an ensemble of methods is proposed as the best approach to perform TI,
	how would this be addressed? Is there an objective means of selecting good performing
	methods?}

\exam{Can TI methods be used for detecting and quantifying technical variation in
	between different samples? Be used as a metric for assessing batch integration?}

\exam{Based on the method presented here, we could “learn” the topology of biological
	networks. What is to your insight the transposability of these topologies?}

\subsection{Chapter 8}
\exam{\nimp{The eight chapter is a manuscript in preparation about the assessment of TI methods. The
		main discussion topic is the lack of a quantitative measure for the scoring of new TI
		methods. This leads to the publication and development of new methods that do not show
		a clear performance improvement compared to other methods. The main topic touches
		upon some sore points that are inherent in bioinformatics. The main thought I have with
		this chapter is that defining clear goals in an emerging technology is difficult and therefore is
		sometimes ignored by fellow bioinformaticians. This because the lack of standardized
		performance metrics, missing public datasets that are well described and an ad hoc problem
		definition. I follow the idea of defining a clear cut approach to benchmarking (as described
		in chapter nine) and I like the proposition that is started in this manuscript for a discussion
		on good TI metrics.}}

\exam{Would approaches like e.g. DREAM challenges be a means of tackling these issues
	to your opinion?}

\exam{Do you expect that all methods which are mainly developed to answer specific
	biological questions would adhere to the same biological problem definition? Is the fact that
	people are trying to answer different question not a prior to define more generalized
	questions?}

### 2019-11-23 defence_internal/pres_v1.tex
update
~

@@

\begin{frame}
	\frametitle{Research questions}
	\begin{center}
		\only<2>{\includegraphics[width=\linewidth]{figures/11a_fundamental.pdf}}%
		\only<3>{\includegraphics[width=\linewidth]{figures/11b_fundamental.pdf}}%
		\only<4>{\includegraphics[width=\linewidth]{figures/11c_fundamental.pdf}}%
		\only<5>{\includegraphics[width=\linewidth]{figures/11d_fundamental.pdf}}%
	\end{center}
\end{frame}

\begin{frame}
\includegraphics[width=\linewidth]{figures/overview.pdf}
\end{frame}

@@

\begin{frame}
	\vfill
	\begin{center}
		\LARGE Trajectory inference
	\end{center}
	\vfill
\end{frame}

@@

\begin{frame}
	\frametitle{A typical trajectory inference analysis}
	\begin{center}
		\only<1>{\includegraphics[width=\linewidth]{figures/4a_trajectoryinference.pdf}}%
		\only<2>{\includegraphics[width=\linewidth]{figures/4b_trajectoryinference.pdf}}%
		\only<3>{\includegraphics[width=\linewidth]{figures/4c_trajectoryinference.pdf}}%
		\only<4>{\includegraphics[width=\linewidth]{figures/4d_trajectoryinference.pdf}}%
		\only<5>{\includegraphics[width=\linewidth]{figures/4e_trajectoryinference.pdf}}%
	\end{center}
\end{frame}

@@

\begin{frame}
	\vfill
	\begin{center}
		\LARGE Network inference
	\end{center}
	\vfill
\end{frame}

\begin{frame}
	\frametitle{Cellular dynamics is (i.a.) driven by gene regulation}
	\begin{center}
		\only<1>{\includegraphics[width=.8\linewidth]{figures/7a_generegulation.pdf}}%
		\only<2>{\includegraphics[width=.8\linewidth]{figures/7b_generegulation.pdf}}%
		\only<3>{\includegraphics[width=.8\linewidth]{figures/7c_generegulation.pdf}}%
		\only<4>{\includegraphics[width=.8\linewidth]{figures/7d_generegulation.pdf}}%
		\only<5>{\includegraphics[width=.8\linewidth]{figures/7e_generegulation.pdf}}%
		\only<6>{\includegraphics[width=.8\linewidth]{figures/7f_generegulation.pdf}}%
	\end{center}
\end{frame}

@@

\begin{frame}
	\frametitle{Case-specific network inference}
	\begin{center}
		\only<1>{\includegraphics[width=\linewidth]{figures/9a_csni_example.pdf}}%
		\only<2>{\includegraphics[width=\linewidth]{figures/9b_csni_example.pdf}}%
		\only<3>{\includegraphics[width=\linewidth]{figures/9c_csni_example.pdf}}%
		\only<4>{\includegraphics[width=\linewidth]{figures/9d_csni_example.pdf}}%
	\end{center}
	\topright{ch6}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%% BENCHMARKING
\begin{frame}
	\vfill
	\begin{center}
		\LARGE Benchmarking
	\end{center}
	\vfill
\end{frame}

\begin{frame}
	\frametitle{Essential guidelines for benchmarking\\computational tools}
	\begin{center}
		\includegraphics[width=\linewidth]{figures/end_part_guidelines.pdf}
	\end{center}
	\topright{ch9}
\end{frame}

@@

\begin{frame}
	\frametitle{Practical implications}
	\begin{itemize}
		\item<2-> Fundamental research in single-cell analyses \\
		  \uncover<3->{\begin{center}\includegraphics[width=.7\linewidth]{figures/11d_fundamental.pdf}\end{center}}
		\item<4-> Benchmarking
		\item<5-> R \& tidyverse
	\end{itemize}
  \topright{all}
\end{frame}

>>>>

\definecolor{set1red}{HTML}{E41A1C}
\definecolor{set1blue}{HTML}{377EB8}
\definecolor{set1green}{HTML}{4DAF4A}

@@

\begin{frame}
	\frametitle{Research questions regarding single-cell biology}
	\begin{center}
		\only<1>{\includegraphics[width=\linewidth]{figures/11a_fundamental.pdf}}%
		\only<2>{\includegraphics[width=\linewidth]{figures/11b_fundamental.pdf}}%
		\only<3>{\includegraphics[width=\linewidth]{figures/11c_fundamental.pdf}}%
		\only<4>{\includegraphics[width=\linewidth]{figures/11d_fundamental.pdf}}%
		\only<5>{\includegraphics[width=\linewidth]{figures/11e_fundamental.pdf}}%
		\only<6>{\includegraphics[width=\linewidth]{figures/11f_fundamental.pdf}}%
	\end{center}
\end{frame}

\begin{frame}
  \only<1>{\includegraphics[width=\linewidth]{figures/overview_categories.pdf}}
  \only<2>{\includegraphics[width=\linewidth]{figures/overview_categories_dots.pdf}}
  \only<3>{\includegraphics[width=\linewidth]{figures/overview.pdf}}
\end{frame}

@@

\begin{frame}
	\vfill
	\begin{center}
		\color{set1blue}
		{\LARGE Trajectory inference} \\
		How do cells change over time?
	\end{center}
	\vfill
\end{frame}

@@

\begin{frame}
	\frametitle{A typical trajectory inference analysis}
	\begin{center}
		\only<1>{\includegraphics[width=\linewidth]{figures/4a_trajectoryinference.pdf}}%
		\only<2>{\includegraphics[width=\linewidth]{figures/4b_trajectoryinference.pdf}}%
		\only<3>{\includegraphics[width=\linewidth]{figures/4c_trajectoryinference.pdf}}%
		\only<4>{\includegraphics[width=\linewidth]{figures/4d_trajectoryinference.pdf}}%
		\only<5>{\includegraphics[width=\linewidth]{figures/4e_trajectoryinference.pdf}}%
		\only<6>{\includegraphics[width=\linewidth]{figures/4f_trajectoryinference.pdf}}%
	\end{center}
\end{frame}

@@

\begin{frame}
	\vfill
	\begin{center}
		\color{set1green}
		{\LARGE Network inference}\\
		What drives cells to change?
	\end{center}
	\vfill
\end{frame}

%\begin{frame}
%	\frametitle{Cellular dynamics is (i.a.) driven by gene regulation}
%	\begin{center}
%		\only<1>{\includegraphics[width=.8\linewidth]{figures/7a_generegulation.pdf}}%
%		\only<2>{\includegraphics[width=.8\linewidth]{figures/7b_generegulation.pdf}}%
%		\only<3>{\includegraphics[width=.8\linewidth]{figures/7c_generegulation.pdf}}%
%		\only<4>{\includegraphics[width=.8\linewidth]{figures/7d_generegulation.pdf}}%
%		\only<5>{\includegraphics[width=.8\linewidth]{figures/7e_generegulation.pdf}}%
%		\only<6>{\includegraphics[width=.8\linewidth]{figures/7f_generegulation.pdf}}%
%	\end{center}
%\end{frame}

@@

\begin{frame}
	\frametitle{Classical vs. sample-specific network inference}
	\begin{center}
		\includegraphics[width=.9\linewidth]{figures/12_nitypes.pdf}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Sample-specific network inference}
	\begin{center}
		\only<1>{\includegraphics[width=\linewidth]{figures/9a_csni_example.pdf}}%
		\only<2>{\includegraphics[width=\linewidth]{figures/9b_csni_example.pdf}}%
		\only<3>{\includegraphics[width=\linewidth]{figures/9c_csni_example.pdf}}%
		\only<4>{\includegraphics[width=\linewidth]{figures/9d_csni_example.pdf}}%
	\end{center}
	\topright{ch6}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%% BENCHMARKING
\begin{frame}
	\vfill
	\begin{center}
		\color{set1blue}
		{\LARGE Benchmarking} \\
		How to quantify the accuracy of computational tools?
	\end{center}
	\vfill
\end{frame}

\begin{frame}
	\frametitle{Essential guidelines for benchmarking\\computational tools}
	\begin{center}
		\includegraphics[width=\linewidth]{figures/end_part_guidelines.pdf}
	\end{center}
	{\scriptsize 
	\begin{flushright}
		In collaboration with the labs of Mark Robinson,\\Anne-Laure Boulesteix, and Paul P. Gardner.
	\end{flushright}}
	\topright{ch9}
\end{frame}

@@

\begin{frame}
	\frametitle{Practical implications}
  \uncover<2->{\begin{center}\includegraphics[width=.7\linewidth]{figures/11f_fundamental.pdf}\end{center}}
	\uncover<3->{Contributions to open-source software in the R ecosystem}
  \topright{all}
\end{frame}
